

The reconstruction of Gene Regulatory Networks from experimental time-series data is a fundamental problem of systems biology, allowing for a deeper understanding of cellular mechanisms. 
Boolean Networks provide a robust framework for modeling these interactions, where the state of each molecular component is represented as a binary value.
However, accurately inferring the underlying logic of these networks from limited and noisy data remains a significant challenge.

This report focuses on benchmarking of \textbf{BNFinder2}, a specialized tool designed for the reconstruction of Dynamic Bayesian Networks and from such networks we will try to reconstruct refernce Boolean Network.
The primary objective of Part I is to establish a performance baseline using synthetic Boolean networks.
With a controlled synthetic environment, we can treat the generated networks as a benchmark  allowing for a quantitative evaluation of the algorithm's reconstruction accuracy that would be impossible with empirical data alone.

Our benchmarking approach explores quite broad hyperparameter space including:
\begin{itemize}
    \item \textbf{Network-dependent properties:} Scaling the dimensionality from 5D to 16D.
    \item \textbf{Trajectory-dependent properties:} Varying trajectory lengths, dataset sizes, and sampling frequencies to simulate different experimental conditions.
    \item \textbf{Dynamics-dependent properties:} Explorint the impact of attractor ratios on the inference process, the balance between steady-state and transient information provided in the trajectories.
\end{itemize}

By finding the optimal parameter combinations, Part I provides the necessary guidelines for the biological validation part.
The insights gained here guide the application of BNFinder2 to real biological models in Part II.